---
title: "Fast Food & Health in America"
subtitle: "R Shiny Dashboard — Full Source Code"
author: "Dashboard Documentation"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: kate
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    df_print: paged
    css: |
      body { font-family: 'DM Sans', sans-serif; }
      h1, h2, h3 { font-family: serif; }
      .section-tag {
        font-size: 0.65rem; letter-spacing: 0.18em; text-transform: uppercase;
        color: #d4380d; font-family: monospace; display: block; margin-bottom: 4px;
      }
      pre { border-left: 4px solid #d4380d; border-radius: 0; }
      .tocify { border-radius: 0; }
      .list-group-item.active { background: #1a1612; border-color: #1a1612; }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  eval    = FALSE,   # set TRUE to run code chunks when knitting
  message = FALSE,
  warning = FALSE,
  comment = ""
)
```

---

> **What this document is:** A fully documented walkthrough of every R file in the
> Fast Food & Health Shiny dashboard. Each section corresponds to one source file.
> Set `eval = TRUE` in the setup chunk (above) to execute the code while knitting.

---

# Project Structure

```
fastfood_shiny/
├── app.R                    ← Entry point
├── install_packages.R       ← Run once to install deps
├── renv.lock                ← Reproducible package versions
├── R/
│   ├── data_load.R          ← Load & clean all 15 CSVs
│   ├── ui_components.R      ← CSS, KPI boxes, data explorer
│   ├── server_chains.R      ← Chains tab module
│   ├── server_nutrition.R   ← Nutrition tab module
│   ├── server_obesity.R     ← Obesity & Poverty tab module
│   ├── server_race.R        ← Demographics tab module
│   └── server_map.R         ← Interactive Leaflet map module
└── data/                    ← All 15 CSV datasets
```

---

# `install_packages.R` — Dependencies {#install}

<span class="section-tag">Run once before launching the app</span>

```{r install-packages}
# ============================================================
#  install_packages.R
#  Run this ONCE to install all dependencies.
#  From R console:  source("install_packages.R")
# ============================================================

pkgs <- c(
  # Core Shiny
  "shiny",
  "bslib",            # Bootstrap 5 theming + modern layout
  "shinycssloaders",  # Loading spinners

  # Data wrangling
  "dplyr",
  "tidyr",
  "readr",
  "stringr",
  "forcats",

  # Visualization
  "plotly",           # Interactive charts
  "ggplot2",
  "scales",
  "viridis",
  "RColorBrewer",

  # Map
  "leaflet",
  "leaflet.extras",   # Heatmap layer + extra plugins
  "geojsonsf",        # Load GeoJSON for state boundaries

  # Tables
  "DT",

  # Reproducibility
  "renv"
)

# Install any missing packages
missing_pkgs <- pkgs[!pkgs %in% installed.packages()[, "Package"]]

if (length(missing_pkgs) > 0) {
  message("Installing: ", paste(missing_pkgs, collapse = ", "))
  install.packages(missing_pkgs, repos = "https://cloud.r-project.org")
} else {
  message("All packages already installed!")
}

# Verify all loaded
loaded <- sapply(pkgs, requireNamespace, quietly = TRUE)
if (all(loaded)) {
  message("\n✓ All packages ready. Run: shiny::runApp()")
} else {
  message("\n✗ Could not load: ", paste(names(loaded[!loaded]), collapse = ", "))
}
```

---

# `app.R` — Entry Point {#app}

<span class="section-tag">The main file — defines the UI layout, theme, and top-level server</span>

```{r app-r}
# ============================================================
#  Fast Food & Health in America — Shiny Dashboard
#  app.R  (single-file entry point; sources modular R files)
# ============================================================

library(shiny)
library(bslib)
library(dplyr)
library(readr)
library(leaflet)
library(leaflet.extras)
library(plotly)
library(DT)
library(ggplot2)
library(scales)
library(tidyr)
library(stringr)
library(viridis)
library(RColorBrewer)

# Source modular helpers
source("R/data_load.R")
source("R/ui_components.R")
source("R/server_chains.R")
source("R/server_nutrition.R")
source("R/server_obesity.R")
source("R/server_race.R")
source("R/server_map.R")

# ── Load data once at startup ────────────────────────────────
dat <- load_all_data()

# ── Theme ────────────────────────────────────────────────────
app_theme <- bs_theme(
  version       = 5,
  bg            = "#faf7f2",
  fg            = "#2d2520",
  primary       = "#d4380d",
  secondary     = "#8c7355",
  success       = "#27ae60",
  info          = "#2980b9",
  warning       = "#e67e22",
  danger        = "#c0392b",
  base_font     = font_google("DM Sans"),
  heading_font  = font_google("DM Serif Display"),
  code_font     = font_google("Space Mono"),
  "navbar-bg"   = "#1a1612",
  "card-bg"     = "#ffffff",
  "body-bg"     = "#faf7f2"
)

# ── UI ───────────────────────────────────────────────────────
ui <- page_navbar(
  title = tags$span(
    tags$span("Fast Food", style = "font-family:'DM Serif Display';"),
    tags$em(" & ", style = "color:#e67e22; font-style:italic;"),
    tags$span("Health",   style = "font-family:'DM Serif Display';")
  ),
  theme    = app_theme,
  bg       = "#1a1612",
  fillable = TRUE,
  id       = "main_nav",

  header = tags$head(
    tags$link(rel = "stylesheet",
      href = "https://fonts.googleapis.com/css2?family=DM+Sans:wght@300;400;500;700&family=DM+Serif+Display:ital@0;1&family=Space+Mono:wght@400;700&display=swap"),
    tags$style(HTML(app_css()))
  ),

  nav_panel("Overview",     icon = icon("chart-bar"),       hero_ui("hero", dat)),
  nav_panel("Chains",       icon = icon("store"),            chains_ui("chains")),
  nav_panel("Nutrition",    icon = icon("apple-alt"),        nutrition_ui("nutrition")),
  nav_panel("Obesity",      icon = icon("heartbeat"),        obesity_ui("obesity")),
  nav_panel("Demographics", icon = icon("users"),            race_ui("race")),
  nav_panel("Map",          icon = icon("map-marked-alt"),   map_ui("map")),
  nav_panel("Data",         icon = icon("table"),            data_explorer_ui("data_exp", dat)),

  nav_spacer(),
  nav_item(
    tags$a(icon("github"), " GitHub",
      href = "https://github.com/YOUR_USERNAME/fastfood-health-dashboard",
      target = "_blank", class = "btn btn-sm btn-outline-light me-2")
  )
)

# ── Server ───────────────────────────────────────────────────
server <- function(input, output, session) {
  chains_server("chains",       dat)
  nutrition_server("nutrition", dat)
  obesity_server("obesity",     dat)
  race_server("race",           dat)
  map_server("map",             dat)
  data_explorer_server("data_exp", dat)

  # Hero KPI outputs
  output$hero_kpi_chains    <- renderText(nrow(dat$chains))
  output$hero_kpi_locations <- renderText(format(nrow(dat$locations), big.mark = ","))
  output$hero_kpi_items     <- renderText(format(nrow(dat$nutrition), big.mark = ","))
  output$hero_kpi_states    <- renderText(n_distinct(dat$obesity$NAME))
}

shinyApp(ui, server)
```

---

# `R/data_load.R` — Data Loading {#data-load}

<span class="section-tag">Reads and cleans every CSV — called once at startup</span>

```{r data-load}
# ============================================================
#  R/data_load.R
#  Loads and cleans every CSV dataset.
#  Returns a named list `dat` consumed throughout the app.
# ============================================================

load_all_data <- function(data_dir = "data") {

  message("Loading datasets...")

  # ── 1. Fast food restaurant locations ──────────────────────
  locations <- read_csv(
    file.path(data_dir, "FastFoodRestaurants.csv"),
    show_col_types = FALSE
  ) |>
    rename_with(str_to_lower) |>
    filter(!is.na(latitude), !is.na(longitude)) |>
    filter(between(longitude, -170, -60),
           between(latitude,   17,   72)) |>
    mutate(
      name  = str_trim(name),
      state = str_trim(province),
      city  = str_trim(city)
    ) |>
    select(name, latitude, longitude, city, state)

  # ── 2. Nutrition per menu item ──────────────────────────────
  nutrition <- read_csv(
    file.path(data_dir, "fastfood.csv"),
    show_col_types = FALSE
  ) |>
    rename_with(str_to_lower) |>
    mutate(across(c(calories, cal_fat, total_fat, sat_fat,
                    trans_fat, cholesterol, sodium,
                    total_carb, fiber, sugar, protein),
                  as.numeric)) |>
    filter(!is.na(calories))

  # ── 3. Top 50 chains ───────────────────────────────────────
  chains <- read_csv(
    file.path(data_dir, "Top_50_Fast-Food_Chains_in_USA.csv"),
    show_col_types = FALSE
  ) |>
    rename(
      chain       = `Fast-Food Chains`,
      sales_m     = `U.S. Systemwide Sales (Millions - U.S Dollars)`,
      avg_unit_k  = `Average Sales per Unit (Thousands - U.S Dollars)`,
      franchised  = `Franchised Stores`,
      company     = `Company Stores`,
      total_units = `2021 Total Units`,
      unit_change = `Total Change in Units from 2020`
    ) |>
    mutate(across(c(sales_m, avg_unit_k, franchised,
                    company, total_units, unit_change), as.numeric)) |>
    arrange(desc(sales_m))

  # ── 4. Obesity by state ────────────────────────────────────
  obesity <- read_csv(
    file.path(data_dir, "National_Obesity_by_State.csv"),
    show_col_types = FALSE
  ) |>
    rename_with(str_to_lower) |>
    rename(NAME = name, obesity_pct = obesity) |>
    filter(!is.na(obesity_pct)) |>
    mutate(obesity_pct = as.numeric(obesity_pct))

  # ── 5. Obesity prevalence county-level ─────────────────────
  obesity_county <- read_csv(
    file.path(data_dir, "Obesity_Prevalence__2023.csv"),
    show_col_types = FALSE
  ) |>
    mutate(value = round(as.numeric(value) * 100, 1)) |>
    filter(!is.na(value))

  # ── 6. Poverty by state (HDPulse) ──────────────────────────
  poverty_raw <- read_csv(
    file.path(data_dir, "HDPulse_data_export-2.csv"),
    skip = 4, show_col_types = FALSE
  )
  colnames(poverty_raw)[1:4] <- c("state", "fips", "poverty_pct", "families_below_poverty")
  poverty <- poverty_raw |>
    filter(!state %in% c("State", "United States", "Puerto Rico", "")) |>
    mutate(poverty_pct = as.numeric(poverty_pct)) |>
    filter(!is.na(poverty_pct)) |>
    select(state, poverty_pct, families_below_poverty)

  # ── 7. Race / ethnicity by state ───────────────────────────
  race_raw_lines <- readLines(file.path(data_dir, "raw_data.csv"))
  race <- read_csv(
    I(race_raw_lines[-(1:2)]),
    show_col_types = FALSE
  ) |>
    rename_with(str_to_lower) |>
    rename(state = location) |>
    select(state, white, black, hispanic, asian,
           `american indian or alaska native`,
           `native hawaiian or pacific islander`,
           `multiple races`) |>
    filter(!state %in% c("", "Location")) |>
    mutate(across(-state, ~ suppressWarnings(as.numeric(.)))) |>
    filter(!is.na(white))

  # ── 8. Historical CPI ──────────────────────────────────────
  cpi <- read_csv(
    file.path(data_dir, "historicalcpi.csv"),
    show_col_types = FALSE
  ) |>
    rename_with(str_to_lower) |>
    rename(item = 1, year = 2, pct_change = 3) |>
    mutate(pct_change = as.numeric(pct_change),
           year       = as.integer(year))

  # ── 9. State population ────────────────────────────────────
  pop <- read_csv(
    file.path(data_dir, "historical_state_population_by_year.csv"),
    col_names = c("state_abbr", "year", "population"),
    show_col_types = FALSE
  ) |>
    mutate(year = as.integer(year), population = as.numeric(population))

  # ── 10. Food access ────────────────────────────────────────
  food_access <- read_csv(
    file.path(data_dir, "food_access.csv"),
    show_col_types = FALSE
  ) |>
    rename_with(str_to_lower) |>
    rename_with(~ str_replace_all(., "\\.", "_"))

  message("All datasets loaded.")

  list(
    locations  = locations,
    nutrition  = nutrition,
    chains     = chains,
    obesity    = obesity,
    obesity_co = obesity_county,
    poverty    = poverty,
    race       = race,
    cpi        = cpi,
    pop        = pop,
    food_access= food_access
  )
}
```

---

# `R/ui_components.R` — Shared UI {#ui-components}

<span class="section-tag">CSS, KPI boxes, section headers, hero panel, data explorer</span>

```{r ui-components}
# ============================================================
#  R/ui_components.R
#  Shared UI helpers: CSS, hero panel, data explorer, KPI boxes
# ============================================================

app_css <- function() {
  "
  body { font-family: 'DM Sans', sans-serif; background:#faf7f2; color:#2d2520; }
  h1,h2,h3,h4,h5 { font-family:'DM Serif Display',serif; font-weight:400; }
  .navbar { border-bottom:3px solid #d4380d !important; }
  .section-label {
    font-family:'Space Mono',monospace; font-size:0.6rem;
    letter-spacing:0.2em; text-transform:uppercase;
    color:#d4380d; margin-bottom:0.2rem;
  }
  .kpi-card {
    background:#1a1612; color:#f0ede6;
    border-radius:0; padding:1.5rem 1.8rem;
    border-top:3px solid #d4380d;
  }
  .kpi-number {
    font-family:'DM Serif Display',serif;
    font-size:2.8rem; line-height:1; color:#e67e22; margin-bottom:0.2rem;
  }
  .kpi-label { font-size:0.72rem; color:rgba(240,237,230,.55);
    text-transform:uppercase; letter-spacing:0.08em; }
  .card { border:1px solid #ede8df; border-radius:0; }
  .card-header { background:#1a1612; color:#f0ede6;
    font-family:'Space Mono',monospace; font-size:0.72rem;
    letter-spacing:0.1em; text-transform:uppercase; }
  table.dataTable thead th { background:#1a1612 !important; color:#f0ede6 !important; }
  .hero-section {
    background:#1a1612; color:#f0ede6;
    padding:3.5rem 2rem 2.5rem; position:relative; overflow:hidden;
  }
  .hero-title { font-size:clamp(2.5rem,6vw,5rem); line-height:0.95;
    letter-spacing:-0.02em; color:#f0ede6; }
  .hero-title em { color:#e67e22; font-style:italic; }
  ::-webkit-scrollbar { width:5px; height:5px; }
  ::-webkit-scrollbar-thumb { background:#d4c9b5; }
  "
}

# ── KPI box ──────────────────────────────────────────────────
kpi_box <- function(output_id, label, icon_name = NULL) {
  div(class = "kpi-card",
    if (!is.null(icon_name))
      tags$div(icon(icon_name), style = "font-size:1.3rem; color:#d4380d; margin-bottom:.4rem;"),
    div(class = "kpi-number", textOutput(output_id, inline = TRUE)),
    div(class = "kpi-label",  label)
  )
}

# ── Section header ───────────────────────────────────────────
section_header <- function(label, title, source_text = NULL) {
  tagList(
    div(class = "section-label", label),
    tags$h4(title, style = "letter-spacing:-0.02em; margin-bottom:0.2rem;"),
    if (!is.null(source_text))
      div(style = "font-family:'Space Mono',monospace; font-size:.65rem;
                   color:#8c7355; margin-bottom:1.2rem;", source_text)
    else
      tags$hr(style = "border-color:#2d2520; margin:.5rem 0 1.2rem;")
  )
}

# ── Hero UI ──────────────────────────────────────────────────
hero_ui <- function(id, dat) {
  ns <- NS(id)
  tagList(
    div(class = "hero-section",
      div(class = "hero-eyebrow",
          "American Fast Food & Public Health · Data Explorer"),
      tags$h1(class = "hero-title",
        "Fast Food,", tags$br(), tags$em("Health & America")),
      p(class = "hero-desc",
        "An interactive deep-dive into fast food chains, nutritional data,
         obesity, poverty, and demographics across the United States —
         powered by 15 real-world datasets.")
    ),
    layout_columns(
      col_widths = c(3,3,3,3), gap = "1px", style = "margin-top:1px;",
      kpi_box("hero_kpi_chains",    "Fast Food Chains",     "store"),
      kpi_box("hero_kpi_locations", "Restaurant Locations", "map-pin"),
      kpi_box("hero_kpi_items",     "Menu Items Analyzed",  "utensils"),
      kpi_box("hero_kpi_states",    "States in Dataset",    "flag-usa")
    ),
    br(),
    layout_columns(
      col_widths = c(4,4,4),
      card(card_header("Highest Obesity State"),
        card_body(p(style="font-family:'DM Serif Display',serif;font-size:2rem;color:#c0392b;margin:0;","Louisiana"),
                  p(style="font-family:'Space Mono',monospace;font-size:.75rem;color:#8c7355;","36.2% adult obesity rate"))),
      card(card_header("Top Chain by Sales"),
        card_body(p(style="font-family:'DM Serif Display',serif;font-size:2rem;color:#e67e22;margin:0;","McDonald's"),
                  p(style="font-family:'Space Mono',monospace;font-size:.75rem;color:#8c7355;","$45.96B systemwide sales · 2021"))),
      card(card_header("Highest Sodium Chain"),
        card_body(p(style="font-family:'DM Serif Display',serif;font-size:2rem;color:#8e44ad;margin:0;","Arby's"),
                  p(style="font-family:'Space Mono',monospace;font-size:.75rem;color:#8c7355;","1,515 mg avg sodium per item")))
    )
  )
}

# ── Data Explorer UI ─────────────────────────────────────────
data_explorer_ui <- function(id, dat) {
  ns <- NS(id)
  tagList(
    br(),
    section_header("Raw Data", "Data Explorer", "Browse & download all datasets"),
    layout_sidebar(
      sidebar = sidebar(width = 220,
        selectInput(ns("dataset"), "Choose Dataset",
          choices = c("Restaurant Locations"="locations","Menu Nutrition"="nutrition",
                      "Top 50 Chains"="chains","Obesity by State"="obesity",
                      "Poverty by State"="poverty","Race / Ethnicity"="race",
                      "Food Access"="food_access","Historical CPI"="cpi")),
        downloadButton(ns("download_csv"), "Download CSV",
          class = "btn-sm w-100",
          style = "border-radius:0;background:#1a1612;color:#f0ede6;border:none;
                   font-family:'Space Mono',monospace;font-size:.7rem;")
      ),
      card(full_screen = TRUE,
           card_body(padding = 0, DTOutput(ns("table"), height = "600px")))
    )
  )
}

data_explorer_server <- function(id, dat) {
  moduleServer(id, function(input, output, session) {
    current_df <- reactive({
      switch(input$dataset,
        locations   = dat$locations,
        nutrition   = dat$nutrition,
        chains      = dat$chains,
        obesity     = dat$obesity,
        poverty     = dat$poverty,
        race        = dat$race,
        food_access = dat$food_access,
        cpi         = dat$cpi
      )
    })
    output$table <- renderDT({
      datatable(current_df(), options = list(pageLength=25, scrollX=TRUE, dom="ftip"),
                class = "compact stripe", rownames = FALSE)
    })
    output$download_csv <- downloadHandler(
      filename = function() paste0(input$dataset, "_fastfood_health.csv"),
      content  = function(file) write_csv(current_df(), file)
    )
  })
}
```

---

# `R/server_chains.R` — Chains Module {#chains}

<span class="section-tag">Sales bar chart · bubble chart · full sortable table</span>

```{r server-chains}
# ============================================================
#  R/server_chains.R  —  Top 50 Chains tab
# ============================================================

chains_ui <- function(id) {
  ns <- NS(id)
  tagList(
    br(),
    section_header("Top 50 Fast Food Chains", "Sales & Market Presence",
                   "Source: QSR Magazine · 2021"),
    tabsetPanel(type = "tabs",

      tabPanel("Sales Bar Chart", br(),
        layout_sidebar(
          sidebar = sidebar(width = 240,
            sliderInput(ns("n_chains"), "Number of chains:", min=5, max=50, value=20, step=5),
            selectInput(ns("sort_by"), "Sort by:",
              choices = c("Sales ($M)"="sales_m","Total Units"="total_units",
                          "Avg Sales/Unit ($K)"="avg_unit_k","Unit Change"="unit_change")),
            hr(),
            radioButtons(ns("color_by"), "Color by:",
              choices = c("Sales"="sales_m","Units"="total_units","Change"="unit_change"))
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("chains_bar"), height="600px")))
        )
      ),

      tabPanel("Bubble Chart", br(),
        layout_sidebar(
          sidebar = sidebar(width = 240,
            selectInput(ns("bub_x"), "X axis:",
              choices = c("Sales ($M)"="sales_m","Total Units"="total_units","Avg Sales/Unit"="avg_unit_k")),
            selectInput(ns("bub_y"), "Y axis:",
              choices = c("Avg Sales/Unit"="avg_unit_k","Sales ($M)"="sales_m","Unit Change"="unit_change")),
            selectInput(ns("bub_size"), "Bubble size:",
              choices = c("Total Units"="total_units","Sales ($M)"="sales_m")),
            checkboxInput(ns("bub_label"), "Show labels", TRUE)
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("chains_bubble"), height="580px")))
        )
      ),

      tabPanel("Full Table", br(),
        card(card_body(padding=0, DTOutput(ns("chains_table"))))
      )
    )
  )
}

chains_server <- function(id, dat) {
  moduleServer(id, function(input, output, session) {

    filtered_chains <- reactive({
      dat$chains |>
        arrange(desc(.data[[input$sort_by]])) |>
        slice_head(n = input$n_chains)
    })

    output$chains_bar <- renderPlotly({
      df <- filtered_chains() |>
        mutate(chain = fct_reorder(chain, .data[[input$sort_by]]))
      color_vals <- df[[input$color_by]]
      pal <- colorRampPalette(c("#f39c12","#e67e22","#d4380d","#922b21"))(100)
      norm  <- (color_vals - min(color_vals, na.rm=TRUE)) /
               (max(color_vals, na.rm=TRUE) - min(color_vals, na.rm=TRUE) + 1e-9)
      colors <- pal[pmax(1, ceiling(norm * 100))]

      plot_ly(df, x=~.data[[input$sort_by]], y=~chain, type="bar",
              orientation="h", marker=list(color=colors),
              hovertemplate="<b>%{y}</b><br>Value: %{x:,.0f}<extra></extra>") |>
        layout(xaxis=list(title=input$sort_by, gridcolor="#ede8df"),
               yaxis=list(title=""),
               plot_bgcolor="#ffffff", paper_bgcolor="#ffffff",
               margin=list(l=10,r=20,t=20,b=40),
               font=list(family="DM Sans", color="#2d2520"))
    })

    output$chains_bubble <- renderPlotly({
      df <- dat$chains
      p <- plot_ly(df, x=~.data[[input$bub_x]], y=~.data[[input$bub_y]],
        size=~.data[[input$bub_size]], text=~chain, type="scatter", mode="markers",
        marker=list(sizemode="area",
          sizeref = 2 * max(df[[input$bub_size]], na.rm=TRUE) / (50^2),
          color=~sales_m,
          colorscale=list(c(0,"#f5cba7"),c(0.5,"#e67e22"),c(1,"#922b21")),
          showscale=TRUE, opacity=0.85),
        hovertemplate="<b>%{text}</b><br>%{x:,.0f} / %{y:,.0f}<extra></extra>")
      if (input$bub_label)
        p <- p |> add_annotations(x=df[[input$bub_x]], y=df[[input$bub_y]],
          text=df$chain, showarrow=FALSE, font=list(size=8,color="#2d2520"),
          xanchor="left", xshift=6)
      p |> layout(plot_bgcolor="#ffffff", paper_bgcolor="#ffffff",
                  font=list(family="DM Sans",color="#2d2520"), showlegend=FALSE)
    })

    output$chains_table <- renderDT({
      dat$chains |>
        mutate(sales_m=dollar(sales_m,suffix="M",prefix="$"),
               avg_unit_k=dollar(avg_unit_k,suffix="K",prefix="$"),
               total_units=comma(total_units),
               unit_change=ifelse(unit_change>=0,paste0("+",comma(unit_change)),comma(unit_change))) |>
        datatable(rownames=FALSE,
                  options=list(pageLength=25,dom="ftip",scrollX=TRUE),
                  class="compact stripe")
    })
  })
}
```

---

# `R/server_nutrition.R` — Nutrition Module {#nutrition}

<span class="section-tag">Per-restaurant averages · heatmap · scatter · violin distributions</span>

```{r server-nutrition}
# ============================================================
#  R/server_nutrition.R  —  Nutrition tab
# ============================================================

nutrition_ui <- function(id) {
  ns <- NS(id)
  tagList(
    br(),
    section_header("Menu Nutrition", "What's Really in Your Meal?",
                   "Source: FDA Menu Labeling Data · 515 items across 8 chains"),
    tabsetPanel(type = "tabs",

      tabPanel("By Restaurant", br(),
        layout_sidebar(
          sidebar = sidebar(width = 240,
            selectInput(ns("nutr_metric"), "Metric:",
              choices = c("Calories"="calories","Sodium (mg)"="sodium",
                          "Total Fat (g)"="total_fat","Protein (g)"="protein",
                          "Sugar (g)"="sugar","Sat. Fat (g)"="sat_fat")),
            radioButtons(ns("nutr_stat"), "Summary statistic:",
              choices=c("Mean"="mean","Median"="median","Max"="max","Min"="min"),
              selected="mean"),
            hr(),
            checkboxInput(ns("nutr_sort"), "Sort by value", TRUE)
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("nutr_bar"), height="450px")))
        )
      ),

      tabPanel("Nutrient Heatmap", br(),
        card(full_screen=TRUE, card_body(plotlyOutput(ns("nutr_heatmap"), height="500px")))
      ),

      tabPanel("Item Scatter", br(),
        layout_sidebar(
          sidebar = sidebar(width = 240,
            selectInput(ns("scatter_x"), "X axis:",
              choices=c("Calories"="calories","Sodium"="sodium","Fat"="total_fat"),
              selected="calories"),
            selectInput(ns("scatter_y"), "Y axis:",
              choices=c("Sodium"="sodium","Calories"="calories","Fat"="total_fat"),
              selected="sodium"),
            selectInput(ns("scatter_color"), "Color by:",
              choices=c("Restaurant"="restaurant","Salad"="salad"))
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("nutr_scatter"), height="550px")))
        )
      ),

      tabPanel("Distributions", br(),
        layout_sidebar(
          sidebar = sidebar(width = 240,
            selectInput(ns("dist_metric"), "Metric:",
              choices=c("Calories"="calories","Sodium"="sodium","Fat"="total_fat","Protein"="protein")),
            selectInput(ns("dist_chain"), "Filter chains:", choices=NULL, multiple=TRUE)
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("nutr_violin"), height="500px")))
        )
      ),

      tabPanel("All Items", br(),
        card(card_body(padding=0, DTOutput(ns("nutr_table"))))
      )
    )
  )
}

nutrition_server <- function(id, dat) {
  moduleServer(id, function(input, output, session) {

    observe({
      chains <- sort(unique(dat$nutrition$restaurant))
      updateSelectInput(session, "dist_chain", choices=chains, selected=chains)
    })

    output$nutr_bar <- renderPlotly({
      stat_fn <- switch(input$nutr_stat,
        mean=~mean(.x,na.rm=TRUE), median=~median(.x,na.rm=TRUE),
        max=~max(.x,na.rm=TRUE),   min=~min(.x,na.rm=TRUE))
      df <- dat$nutrition |> group_by(restaurant) |>
        summarise(value=stat_fn(!!sym(input$nutr_metric)), n=n(), .groups="drop")
      if (input$nutr_sort) df <- df |> arrange(desc(value))
      df <- df |> mutate(restaurant=fct_inorder(restaurant))
      colors <- colorRampPalette(c("#f5cba7","#e67e22","#d4380d","#922b21"))(nrow(df))
      plot_ly(df, x=~value, y=~restaurant, type="bar", orientation="h",
              marker=list(color=colors),
              hovertemplate="<b>%{y}</b><br>Value: %{x:.1f} (n=%{customdata})<extra></extra>",
              customdata=~n) |>
        layout(xaxis=list(title=input$nutr_metric, gridcolor="#ede8df"),
               yaxis=list(title=""), plot_bgcolor="#ffffff", paper_bgcolor="#ffffff",
               margin=list(l=10,r=50,t=10,b=40), font=list(family="DM Sans",color="#2d2520"))
    })

    output$nutr_heatmap <- renderPlotly({
      metrics <- c("calories","sodium","total_fat","protein","sugar","sat_fat","cholesterol")
      df <- dat$nutrition |> group_by(restaurant) |>
        summarise(across(all_of(metrics), ~mean(.x,na.rm=TRUE)), .groups="drop")
      mat <- df |> select(all_of(metrics)) |>
        mutate(across(everything(),
          ~(.x-min(.x,na.rm=TRUE))/(max(.x,na.rm=TRUE)-min(.x,na.rm=TRUE)+1e-9)))
      plot_ly(x=metrics, y=df$restaurant,
              z=matrix(unlist(lapply(seq_len(nrow(mat)),function(i)as.numeric(mat[i,]))),
                       nrow=nrow(df),byrow=TRUE),
              type="heatmap",
              colorscale=list(c(0,"#f9f3e8"),c(0.5,"#e67e22"),c(1,"#7b241c")),
              hovertemplate="<b>%{y}</b> — %{x}: %{z:.2f}<extra></extra>") |>
        layout(xaxis=list(title="",tickangle=-20), yaxis=list(title=""),
               plot_bgcolor="#ffffff", paper_bgcolor="#ffffff",
               margin=list(l=10,r=80,t=10,b=60), font=list(family="DM Sans",color="#2d2520"))
    })

    output$nutr_scatter <- renderPlotly({
      df <- dat$nutrition |>
        filter(!is.na(.data[[input$scatter_x]]), !is.na(.data[[input$scatter_y]]))
      pal <- c("#e8a000","#e63000","#7b2d8b","#c0392b","#8b2500","#00437a","#009a44","#e67e22")
      plot_ly(df, x=~.data[[input$scatter_x]], y=~.data[[input$scatter_y]],
              color=~.data[[input$scatter_color]], colors=pal,
              type="scatter", mode="markers", text=~item,
              marker=list(size=7,opacity=0.75),
              hovertemplate="<b>%{text}</b><br>%{x:.0f} / %{y:.0f}<extra></extra>") |>
        layout(xaxis=list(title=input$scatter_x,gridcolor="#ede8df"),
               yaxis=list(title=input$scatter_y,gridcolor="#ede8df"),
               plot_bgcolor="#ffffff", paper_bgcolor="#ffffff",
               font=list(family="DM Sans",color="#2d2520"))
    })

    output$nutr_violin <- renderPlotly({
      df <- dat$nutrition |> filter(restaurant %in% input$dist_chain,
                                    !is.na(.data[[input$dist_metric]]))
      pal <- c("#e8a000","#e63000","#7b2d8b","#c0392b","#8b2500","#00437a","#009a44","#e67e22")
      plot_ly(df, x=~restaurant, y=~.data[[input$dist_metric]],
              color=~restaurant, colors=pal, type="violin",
              box=list(visible=TRUE), meanline=list(visible=TRUE,color="white",width=2),
              hovertemplate="<b>%{x}</b><br>%{y:.0f}<extra></extra>") |>
        layout(xaxis=list(title=""), yaxis=list(title=input$dist_metric,gridcolor="#ede8df"),
               plot_bgcolor="#ffffff", paper_bgcolor="#ffffff",
               font=list(family="DM Sans",color="#2d2520"), showlegend=FALSE)
    })

    output$nutr_table <- renderDT({
      datatable(dat$nutrition, rownames=FALSE, filter="top",
                options=list(pageLength=20,dom="ftip",scrollX=TRUE),
                class="compact stripe")
    })
  })
}
```

---

# `R/server_obesity.R` — Obesity & Poverty Module {#obesity}

<span class="section-tag">State rankings · lollipop chart · scatter · deep-dive gauge dials</span>

```{r server-obesity}
# ============================================================
#  R/server_obesity.R  —  Obesity & Poverty tab
# ============================================================

obesity_ui <- function(id) {
  ns <- NS(id)
  tagList(
    br(),
    section_header("State Health Data", "Obesity & Poverty Across America",
                   "Sources: CDC BRFSS 2023 · U.S. Census SAIPE 2019–2023"),
    tabsetPanel(type = "tabs",

      tabPanel("Obesity Ranking", br(),
        layout_columns(col_widths=c(7,5),
          card(full_screen=TRUE, card_header("All States — Obesity Rate (%)"),
               card_body(plotlyOutput(ns("obesity_lollipop"), height="700px"))),
          card(full_screen=TRUE, card_header("Summary"),
               card_body(plotlyOutput(ns("obesity_hist"),height="280px"),
                         hr(),
                         plotlyOutput(ns("obesity_box"),height="280px")))
        )
      ),

      tabPanel("Poverty Ranking", br(),
        layout_columns(col_widths=c(7,5),
          card(full_screen=TRUE, card_header("States — Poverty Rate (%)"),
               card_body(plotlyOutput(ns("poverty_bar"), height="600px"))),
          card(full_screen=TRUE, card_header("Poverty vs Obesity Scatter"),
               card_body(plotlyOutput(ns("pov_ob_scatter"), height="560px")))
        )
      ),

      tabPanel("State Deep-Dive", br(),
        layout_sidebar(
          sidebar = sidebar(width=240,
            selectInput(ns("state_select"), "Choose State:", choices=NULL, selected="Louisiana")),
          layout_columns(col_widths=c(6,6),
            card(card_header("Health Metrics"),
                 card_body(plotlyOutput(ns("state_gauge"), height="320px"))),
            card(card_header("Peer Comparison"),
                 card_body(plotlyOutput(ns("state_peer"), height="320px")))
          )
        )
      )
    )
  )
}

obesity_server <- function(id, dat) {
  moduleServer(id, function(input, output, session) {

    ob_pov <- reactive({
      dat$obesity |> left_join(dat$poverty, by=c("NAME"="state")) |>
        filter(!NAME %in% c("Puerto Rico", NA))
    })

    observe({
      updateSelectInput(session, "state_select",
        choices=sort(ob_pov()$NAME), selected="Louisiana")
    })

    output$obesity_lollipop <- renderPlotly({
      df <- ob_pov() |> arrange(desc(obesity_pct)) |>
        mutate(NAME=fct_inorder(NAME),
               color=case_when(obesity_pct>=34~"#922b21",obesity_pct>=30~"#d4380d",
                               obesity_pct>=26~"#e67e22",TRUE~"#27ae60"))
      plot_ly(df) |>
        add_segments(x=0,xend=~obesity_pct,y=~NAME,yend=~NAME,
                     line=list(color="rgba(0,0,0,.1)",width=1)) |>
        add_markers(x=~obesity_pct,y=~NAME,
                    marker=list(color=~color,size=10,line=list(color="white",width=1)),
                    text=~paste0(NAME,": ",obesity_pct,"%"), hoverinfo="text") |>
        layout(xaxis=list(title="Adult Obesity Rate (%)",range=c(18,40),gridcolor="#ede8df"),
               yaxis=list(title="",tickfont=list(size=10)),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               margin=list(l=10,r=10,t=10,b=40),
               font=list(family="DM Sans",color="#2d2520"),showlegend=FALSE)
    })

    output$poverty_bar <- renderPlotly({
      df <- ob_pov() |> filter(!is.na(poverty_pct)) |> arrange(desc(poverty_pct)) |>
        mutate(NAME=fct_inorder(NAME),
               color=case_when(poverty_pct>=13~"#6c3483",poverty_pct>=11~"#8e44ad",
                               poverty_pct>=9~"#2c3e50",poverty_pct>=7~"#2980b9",TRUE~"#16a085"))
      plot_ly(df,x=~poverty_pct,y=~NAME,type="bar",orientation="h",
              marker=list(color=~color),
              hovertemplate="<b>%{y}</b><br>Poverty: %{x:.1f}%<extra></extra>") |>
        layout(xaxis=list(title="Poverty Rate (%)",gridcolor="#ede8df"),
               yaxis=list(title="",tickfont=list(size=10)),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               margin=list(l=10,r=10,t=10,b=40),font=list(family="DM Sans",color="#2d2520"))
    })

    output$pov_ob_scatter <- renderPlotly({
      df <- ob_pov() |> filter(!is.na(poverty_pct),!is.na(obesity_pct)) |>
        mutate(color=case_when(obesity_pct>=34~"#922b21",obesity_pct>=30~"#d4380d",
                               obesity_pct>=26~"#e67e22",TRUE~"#27ae60"))
      cor_val <- round(cor(df$poverty_pct,df$obesity_pct,use="complete.obs"),2)
      plot_ly(df,x=~poverty_pct,y=~obesity_pct,type="scatter",mode="markers+text",
              text=~NAME,textposition="top right",textfont=list(size=8,color="#8c7355"),
              marker=list(color=~color,size=10,opacity=0.8),
              hovertemplate="<b>%{text}</b><br>Poverty: %{x:.1f}%<br>Obesity: %{y:.1f}%<extra></extra>") |>
        layout(title=list(text=paste0("Pearson r = ",cor_val),font=list(size=11,color="#8c7355"),x=0.01),
               xaxis=list(title="Poverty Rate (%)",gridcolor="#ede8df"),
               yaxis=list(title="Obesity Rate (%)",gridcolor="#ede8df"),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",showlegend=FALSE,
               margin=list(l=20,r=20,t=30,b=50),font=list(family="DM Sans",color="#2d2520"))
    })

    output$state_gauge <- renderPlotly({
      req(input$state_select)
      df  <- ob_pov() |> filter(NAME==input$state_select)
      avg_ob  <- mean(ob_pov()$obesity_pct,  na.rm=TRUE)
      avg_pov <- mean(ob_pov()$poverty_pct,  na.rm=TRUE)
      plot_ly() |>
        add_trace(type="indicator", mode="gauge+number+delta",
          value=df$obesity_pct[1],
          title=list(text="Obesity Rate (%)", font=list(size=13)),
          delta=list(reference=avg_ob,valueformat=".1f"),
          gauge=list(axis=list(range=list(18,40)),bar=list(color="#d4380d"),
            threshold=list(line=list(color="#2d2520",width=2),thickness=0.75,value=avg_ob)),
          domain=list(x=c(0,1),y=c(0.5,1))) |>
        add_trace(type="indicator", mode="gauge+number+delta",
          value=ifelse(!is.na(df$poverty_pct[1]),df$poverty_pct[1],0),
          title=list(text="Poverty Rate (%)", font=list(size=13)),
          delta=list(reference=avg_pov,valueformat=".1f"),
          gauge=list(axis=list(range=list(0,16)),bar=list(color="#8e44ad"),
            threshold=list(line=list(color="#2d2520",width=2),thickness=0.75,value=avg_pov)),
          domain=list(x=c(0,1),y=c(0,0.45))) |>
        layout(plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               font=list(family="DM Sans",color="#2d2520"))
    })

    output$obesity_hist <- renderPlotly({
      df <- ob_pov() |> filter(!is.na(obesity_pct))
      plot_ly(df,x=~obesity_pct,type="histogram",nbinsx=14,
              marker=list(color="#d4380d",line=list(color="white",width=0.8))) |>
        layout(xaxis=list(title="Obesity Rate (%)",gridcolor="#ede8df"),
               yaxis=list(title="# States",gridcolor="#ede8df"),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               margin=list(l=10,r=10,t=5,b=40),font=list(family="DM Sans",color="#2d2520"))
    })

    output$obesity_box <- renderPlotly({
      df <- ob_pov() |> filter(!is.na(obesity_pct))
      plot_ly(df,x=~obesity_pct,type="box",
              marker=list(color="#d4380d",size=4), line=list(color="#922b21"),
              fillcolor="rgba(212,56,13,0.15)") |>
        layout(xaxis=list(title="Obesity Rate (%)",gridcolor="#ede8df"),
               yaxis=list(showticklabels=FALSE),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               margin=list(l=10,r=10,t=5,b=40),font=list(family="DM Sans",color="#2d2520"))
    })

    output$state_peer <- renderPlotly({
      req(input$state_select)
      df <- ob_pov() |> filter(!is.na(obesity_pct)) |>
        arrange(desc(obesity_pct)) |>
        mutate(color=ifelse(NAME==input$state_select,"#d4380d","#d4c9b5")) |>
        slice_head(n=20)
      plot_ly(df,x=~obesity_pct,y=~fct_inorder(NAME),type="bar",orientation="h",
              marker=list(color=~color),
              hovertemplate="<b>%{y}</b>: %{x:.1f}%<extra></extra>") |>
        layout(xaxis=list(title="Obesity Rate (%)",gridcolor="#ede8df"),
               yaxis=list(title="",tickfont=list(size=10)),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               margin=list(l=10,r=10,t=30,b=40),font=list(family="DM Sans",color="#2d2520"))
    })
  })
}
```

---

# `R/server_race.R` — Demographics Module {#race}

<span class="section-tag">National pie · stacked state bars · heatmap · race × obesity scatter</span>

```{r server-race}
# ============================================================
#  R/server_race.R  —  Demographics tab
# ============================================================

race_ui <- function(id) {
  ns <- NS(id)
  tagList(
    br(),
    section_header("Demographics", "Race & Ethnicity Across States",
                   "Source: KFF State Health Facts · 2024"),
    tabsetPanel(type = "tabs",
      tabPanel("National Overview", br(),
        layout_columns(col_widths=c(5,7),
          card(card_header("U.S. Racial Composition"),
               card_body(plotlyOutput(ns("us_pie"), height="400px"))),
          card(full_screen=TRUE, card_header("Fast Food Locations by State (top 20)"),
               card_body(plotlyOutput(ns("ff_state_bar"), height="400px")))
        )
      ),
      tabPanel("Stacked State Bars", br(),
        layout_sidebar(
          sidebar = sidebar(width=220,
            checkboxGroupInput(ns("race_groups"), "Show groups:",
              choices=c("White"="white","Black"="black","Hispanic"="hispanic","Asian"="asian"),
              selected=c("white","black","hispanic","asian")),
            checkboxInput(ns("race_sort"), "Sort by White %", FALSE)
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("race_stack"), height="650px")))
        )
      ),
      tabPanel("Heatmap", br(),
        card(full_screen=TRUE, card_body(plotlyOutput(ns("race_heatmap"), height="600px")))
      ),
      tabPanel("Obesity × Race", br(),
        layout_sidebar(
          sidebar = sidebar(width=220,
            selectInput(ns("race_x"), "Race group:",
              choices=c("White"="white","Black"="black","Hispanic"="hispanic","Asian"="asian")),
            checkboxInput(ns("race_ob_labels"), "Label states", FALSE)
          ),
          card(full_screen=TRUE, card_body(plotlyOutput(ns("race_ob_scatter"), height="520px")))
        )
      )
    )
  )
}

race_server <- function(id, dat) {
  moduleServer(id, function(input, output, session) {
    race_colors <- c(white="#4e9af1",black="#e87040",hispanic="#54c066",asian="#f0c040",
                     `american indian or alaska native`="#c084fc",
                     `native hawaiian or pacific islander`="#fb923c",
                     `multiple races`="#94a3b8")

    output$us_pie <- renderPlotly({
      us     <- dat$race |> filter(state=="United States")
      groups <- c("white","black","hispanic","asian",
                  "american indian or alaska native","multiple races")
      vals   <- as.numeric(us[1, groups])
      labels <- c("White","Black","Hispanic","Asian","Am. Indian","Multiple")
      plot_ly(labels=labels, values=vals, type="pie",
              marker=list(colors=unname(race_colors[groups]),
                          line=list(color="white",width=2)),
              textinfo="label+percent") |>
        layout(showlegend=FALSE,plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               font=list(family="DM Sans",color="#2d2520",size=11))
    })

    output$ff_state_bar <- renderPlotly({
      df <- dat$locations |> count(state,name="n") |>
        arrange(desc(n)) |> slice_head(n=20) |> mutate(state=fct_inorder(state))
      plot_ly(df,x=~n,y=~state,type="bar",orientation="h",
              marker=list(color=colorRampPalette(c("#f5cba7","#d4380d"))(20)),
              hovertemplate="<b>%{y}</b>: %{x:,} locations<extra></extra>") |>
        layout(xaxis=list(title="Number of Locations",gridcolor="#ede8df"),
               yaxis=list(title=""),plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               font=list(family="DM Sans",color="#2d2520"))
    })

    output$race_stack <- renderPlotly({
      df <- dat$race |> filter(state!="United States",!is.na(white)) |>
        select(state,white,black,hispanic,asian)
      if (input$race_sort) df <- df |> arrange(white)
      df <- df |> mutate(state=fct_inorder(state))
      groups <- intersect(input$race_groups, c("white","black","hispanic","asian"))
      labels <- c(white="White",black="Black",hispanic="Hispanic",asian="Asian")
      colors <- c(white="#4e9af1",black="#e87040",hispanic="#54c066",asian="#f0c040")
      p <- plot_ly()
      for (g in groups) {
        p <- p |> add_trace(x=df[[g]]*100, y=~df$state, type="bar",
          orientation="h", name=labels[g], marker=list(color=colors[g]),
          hovertemplate=paste0("<b>%{y}</b><br>",labels[g],": %{x:.1f}%<extra></extra>"))
      }
      p |> layout(barmode="stack",
                  xaxis=list(title="Population Share (%)"),yaxis=list(title=""),
                  legend=list(orientation="h",y=-0.05),
                  plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
                  font=list(family="DM Sans",color="#2d2520"))
    })

    output$race_heatmap <- renderPlotly({
      groups <- c("white","black","hispanic","asian","american indian or alaska native","multiple races")
      df <- dat$race |> filter(state!="United States",!is.na(white)) |>
        arrange(desc(white)) |> mutate(state=fct_inorder(state))
      mat <- df |> select(all_of(groups)) |> mutate(across(everything(),~.x*100))
      plot_ly(x=c("White","Black","Hispanic","Asian","Am. Indian","Multiple"),
              y=df$state, z=as.matrix(mat), type="heatmap",
              colorscale=list(c(0,"#f9f3e8"),c(0.5,"#4e9af1"),c(1,"#1a1612")),
              hovertemplate="<b>%{y}</b> — %{x}: %{z:.1f}%<extra></extra>") |>
        layout(xaxis=list(title="",side="top"),yaxis=list(title="",tickfont=list(size=9)),
               plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",
               margin=list(l=10,r=80,t=40,b=10),font=list(family="DM Sans",color="#2d2520"))
    })

    output$race_ob_scatter <- renderPlotly({
      req(input$race_x)
      df <- dat$race |> filter(state!="United States") |>
        inner_join(dat$obesity,by=c("state"="NAME")) |>
        filter(!is.na(.data[[input$race_x]]),!is.na(obesity_pct)) |>
        mutate(race_pct=.data[[input$race_x]]*100)
      cor_val <- round(cor(df$race_pct,df$obesity_pct,use="complete.obs"),2)
      p <- plot_ly(df,x=~race_pct,y=~obesity_pct,type="scatter",mode="markers",
                   text=~state,
                   marker=list(color=race_colors[input$race_x],size=10,opacity=0.8),
                   hovertemplate="<b>%{text}</b><br>Race: %{x:.1f}%<br>Obesity: %{y:.1f}%<extra></extra>")
      if (input$race_ob_labels)
        p <- p |> add_annotations(x=df$race_pct,y=df$obesity_pct,text=df$state,
          showarrow=FALSE,font=list(size=8,color="#8c7355"),xanchor="left",xshift=5)
      p |> layout(
        title=list(text=paste0("Pearson r = ",cor_val),font=list(size=11,color="#8c7355"),x=0.01),
        xaxis=list(title=paste0(str_to_title(input$race_x)," Population (%)"),gridcolor="#ede8df"),
        yaxis=list(title="Obesity Rate (%)",gridcolor="#ede8df"),
        plot_bgcolor="#ffffff",paper_bgcolor="#ffffff",showlegend=FALSE,
        font=list(family="DM Sans",color="#2d2520"))
    })
  })
}
```

---

# `R/server_map.R` — Interactive Map Module {#map}

<span class="section-tag">9,950 Leaflet markers · cluster toggle · obesity/poverty overlay · click-to-inspect states</span>

```{r server-map}
# ============================================================
#  R/server_map.R  —  Interactive Leaflet Map tab
# ============================================================

CHAIN_COLORS <- c(
  "McDonald's"="# e8a000","Burger King"="#e63000","Taco Bell"="#7b2d8b",
  "Wendy's"="#c0392b","Arby's"="#8b2500","KFC"="#c0392b",
  "Domino's Pizza"="#00437a","Subway"="#009a44","SONIC Drive In"="#0062a3",
  "Hardee's"="#e87722","Jack in the Box"="#e05206","Chick-fil-A"="#dd0031",
  "Five Guys"="#cc0000","Dairy Queen"="#cc0099","Whataburger"="#f5821f"
)

TOP_CHAINS_LIST <- c("McDonald's","Burger King","Taco Bell","Wendy's","Arby's",
                     "KFC","Domino's Pizza","Subway","SONIC Drive In",
                     "Hardee's","Jack in the Box","Chick-fil-A")

get_chain_color <- function(name) {
  col <- CHAIN_COLORS[name]
  ifelse(is.na(col),"#888888",col)
}

map_ui <- function(id) {
  ns <- NS(id)
  tagList(
    div(style="padding:1.5rem 1rem 0;",
      section_header("Interactive Map","9,950 Restaurants Across America",
        "Toggle chains · Overlay obesity or poverty · Click markers for details")),
    div(class="map-controls", style="padding:0 1rem; display:flex; flex-wrap:wrap; gap:.6rem;",
      uiOutput(ns("chain_buttons")),
      div(style="margin-left:auto; display:flex; gap:.8rem; align-items:center;",
        selectInput(ns("overlay"), NULL,
          choices=c("No Overlay"="none","Obesity Rate"="obesity","Poverty Rate"="poverty"),
          width="160px"),
        selectInput(ns("cluster"), NULL,
          choices=c("Cluster markers"="cluster","Show all markers"="all"),
          width="160px"),
        div(style="font-family:'Space Mono',monospace;font-size:.7rem;color:#8c7355;
                   padding:.35rem .8rem;border:1px solid #ede8df;background:white;",
          textOutput(ns("marker_count"), inline=TRUE))
      )
    ),
    div(style="padding:0 1rem;", leafletOutput(ns("leaflet"), height="580px")),
    div(style="padding:.5rem 1rem;", uiOutput(ns("map_legend"))),
    div(style="padding:0 1rem 2rem;", uiOutput(ns("state_info")))
  )
}

map_server <- function(id, dat) {
  moduleServer(id, function(input, output, session) {
    ns <- session$ns

    active_chains <- reactiveVal(TOP_CHAINS_LIST)
    clicked_state <- reactiveVal(NULL)

    norm_name <- function(n) {
      n <- str_replace(n,"^McDonalds$","McDonald's")
      n <- str_replace(n,"^Sonic Drive-In$","SONIC Drive In")
      n
    }

    locations_clean <- reactive({
      dat$locations |> mutate(name=norm_name(name))
    })

    filtered_locs <- reactive({
      locs   <- locations_clean()
      chains <- active_chains()
      if (length(chains) > 0) filter(locs, name %in% chains) else locs
    })

    state_data <- reactive({
      dat$obesity |> left_join(dat$poverty, by=c("NAME"="state"))
    })

    # ── Chain toggle buttons ─────────────────────────────────
    output$chain_buttons <- renderUI({
      tagList(
        lapply(TOP_CHAINS_LIST, function(chain) {
          col       <- get_chain_color(chain)
          is_active <- chain %in% active_chains()
          actionButton(
            inputId = ns(paste0("chain_",make.names(chain))),
            label   = chain,
            style   = paste0(
              "font-size:.6rem;letter-spacing:.06em;text-transform:uppercase;",
              "padding:.2rem .55rem;border-radius:2px;font-weight:700;margin:2px;",
              "font-family:'Space Mono',monospace;border:1.5px solid ",col,";",
              if (is_active) paste0("background:",col,"22;color:",col,";")
              else "background:transparent;color:#bbb;border-color:#ddd;"
            ),
            onclick = sprintf("Shiny.setInputValue('%s','%s',{priority:'event'})",
                              ns("chain_toggle"), chain)
          )
        }),
        actionButton(ns("toggle_all"), "All / None",
          style="font-size:.6rem;letter-spacing:.06em;text-transform:uppercase;
                 padding:.2rem .55rem;border-radius:2px;margin:2px;
                 border:1px solid #8c7355;color:#8c7355;background:transparent;
                 font-family:'Space Mono',monospace;")
      )
    })

    observeEvent(input$chain_toggle, {
      chain   <- input$chain_toggle
      current <- active_chains()
      if (chain %in% current) active_chains(setdiff(current, chain))
      else                     active_chains(union(current, chain))
    })

    observeEvent(input$toggle_all, {
      if (length(active_chains()) > 0) active_chains(character(0))
      else                             active_chains(TOP_CHAINS_LIST)
    })

    output$marker_count <- renderText({
      paste0(format(nrow(filtered_locs()), big.mark=","), " restaurants shown")
    })

    # ── Base map ─────────────────────────────────────────────
    output$leaflet <- renderLeaflet({
      leaflet(options=leafletOptions(minZoom=3)) |>
        addProviderTiles(providers$CartoDB.Positron,
          options=tileOptions(maxZoom=18)) |>
        setView(lng=-96, lat=38.5, zoom=4) |>
        addScaleBar(position="bottomleft")
    })

    # ── Update markers ───────────────────────────────────────
    observeEvent(list(filtered_locs(), input$cluster), {
      locs  <- filtered_locs() |>
        mutate(color=get_chain_color(name),
               popup=paste0("<div style='font-family:DM Sans,sans-serif;'>",
                            "<strong>",name,"</strong><br>",
                            "<span style='color:#8c7355;'>",city,
                            ifelse(state!="",paste0(", ",state),""),"</span></div>"))
      proxy <- leafletProxy(ns("leaflet"), session) |> clearMarkers() |> clearMarkerClusters()
      if (nrow(locs)==0) return()

      if (input$cluster=="cluster") {
        for (chain in unique(locs$name)) {
          sub <- locs |> filter(name==chain)
          col <- get_chain_color(chain)
          proxy <- proxy |> addCircleMarkers(
            data=sub, lng=~longitude, lat=~latitude, popup=~popup,
            radius=5, fillColor=col, fillOpacity=0.85,
            color="rgba(0,0,0,0.2)", weight=0.5,
            clusterOptions=markerClusterOptions(
              iconCreateFunction=JS(sprintf("
                function(cluster) {
                  var n    = cluster.getChildCount();
                  var size = n > 100 ? 44 : n > 20 ? 34 : 26;
                  return L.divIcon({
                    html: '<div style=\"background:%s;color:white;border-radius:50%%;' +
                          'width:'+size+'px;height:'+size+'px;display:flex;' +
                          'align-items:center;justify-content:center;' +
                          'font-family:Space Mono,monospace;font-size:10px;font-weight:700;' +
                          'border:2px solid white;box-shadow:0 1px 4px rgba(0,0,0,.3);\">'+n+'</div>',
                    className:'', iconSize: L.point(size,size)
                  });
                }", col))
            )
          )
        }
      } else {
        proxy |> addCircleMarkers(
          data=locs, lng=~longitude, lat=~latitude, popup=~popup,
          radius=4, fillColor=~color, fillOpacity=0.8,
          color="rgba(0,0,0,0.15)", weight=0.5)
      }
    })

    # ── State overlay ────────────────────────────────────────
    observeEvent(input$overlay, {
      proxy <- leafletProxy(ns("leaflet"),session) |> clearShapes()
      if (input$overlay=="none") return()
      tryCatch({
        states_geo <- geojsonsf::geojson_sf(
          "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json")
        merged <- states_geo |> left_join(state_data(), by=c("name"="NAME"))
        if (input$overlay=="obesity") {
          pal <- colorNumeric(c("#27ae60","#f39c12","#e67e22","#d4380d","#922b21"),
                              domain=c(18,40), na.color="transparent")
          proxy |>
            addPolygons(data=merged, fillColor=~pal(obesity_pct), fillOpacity=0.35,
              color="rgba(0,0,0,.15)", weight=1, layerId=~name,
              label=~paste0(name,": ",obesity_pct,"% obesity"),
              highlightOptions=highlightOptions(weight=2,color="#1a1612",
                fillOpacity=0.55,bringToFront=TRUE)) |>
            addLegend("bottomright",pal=pal,values=merged$obesity_pct,
              title="Obesity Rate (%)",opacity=0.9)
        } else {
          pal <- colorNumeric(c("#16a085","#2980b9","#2c3e50","#8e44ad","#6c3483"),
                              domain=c(5,16), na.color="transparent")
          proxy |>
            addPolygons(data=merged, fillColor=~pal(poverty_pct), fillOpacity=0.35,
              color="rgba(0,0,0,.15)", weight=1, layerId=~name,
              label=~paste0(name,": ",poverty_pct,"% poverty"),
              highlightOptions=highlightOptions(weight=2,color="#1a1612",
                fillOpacity=0.55,bringToFront=TRUE)) |>
            addLegend("bottomright",pal=pal,values=merged$poverty_pct,
              title="Poverty Rate (%)",opacity=0.9)
        }
      }, error=function(e) {
        showNotification("Install 'geojsonsf' for overlays: install.packages('geojsonsf')",
                         type="warning", duration=6)
      })
    })

    observeEvent(input$leaflet_shape_click, {
      clicked_state(input$leaflet_shape_click$id)
    })

    # ── State info panel ─────────────────────────────────────
    output$state_info <- renderUI({
      state_name <- clicked_state()
      if (is.null(state_name)) return(NULL)
      sd <- state_data() |> filter(NAME==state_name)
      if (nrow(sd)==0) return(NULL)
      abbr_map <- c(Alabama="AL",Alaska="AK",Arizona="AZ",Arkansas="AR",California="CA",
        Colorado="CO",Connecticut="CT",Delaware="DE",Florida="FL",Georgia="GA",
        Hawaii="HI",Idaho="ID",Illinois="IL",Indiana="IN",Iowa="IA",Kansas="KS",
        Kentucky="KY",Louisiana="LA",Maine="ME",Maryland="MD",Massachusetts="MA",
        Michigan="MI",Minnesota="MN",Mississippi="MS",Missouri="MO",Montana="MT",
        Nebraska="NE",Nevada="NV",`New Hampshire`="NH",`New Jersey`="NJ",
        `New Mexico`="NM",`New York`="NY",`North Carolina`="NC",`North Dakota`="ND",
        Ohio="OH",Oklahoma="OK",Oregon="OR",Pennsylvania="PA",`Rhode Island`="RI",
        `South Carolina`="SC",`South Dakota`="SD",Tennessee="TN",Texas="TX",
        Utah="UT",Vermont="VT",Virginia="VA",Washington="WA",`West Virginia`="WV",
        Wisconsin="WI",Wyoming="WY",`District of Columbia`="DC")
      abbr       <- abbr_map[state_name]
      state_locs <- locations_clean() |> filter(state %in% c(abbr, state_name))
      top_chains <- state_locs |> count(name,sort=TRUE) |> slice_head(n=5)
      div(style="border:1px solid #ede8df;border-left:4px solid #d4380d;
                 background:white;padding:1.2rem 1.5rem;margin-top:.5rem;",
        div(style="display:flex;gap:2.5rem;flex-wrap:wrap;",
          div(
            tags$h5(paste0("\U0001F4CD ",state_name),
                    style="font-family:'DM Serif Display',serif;font-size:1.4rem;"),
            if (!is.na(sd$obesity_pct))
              p(tags$strong(paste0(sd$obesity_pct,"%"),
                style="font-family:'Space Mono',monospace;"), " adult obesity rate",
                style="font-size:.82rem;margin-bottom:.2rem;"),
            if (!is.na(sd$poverty_pct))
              p(tags$strong(paste0(sd$poverty_pct,"%"),
                style="font-family:'Space Mono',monospace;"), " poverty rate",
                style="font-size:.82rem;")
          ),
          div(
            p("Top chains in dataset:",
              style="font-family:'Space Mono',monospace;font-size:.65rem;color:#8c7355;
                     text-transform:uppercase;letter-spacing:.1em;margin-bottom:.5rem;"),
            tagList(lapply(seq_len(nrow(top_chains)), function(i) {
              chain <- top_chains$name[i]
              col   <- get_chain_color(chain)
              p(tags$span(style=paste0("display:inline-block;width:8px;height:8px;",
                  "border-radius:50%;background:",col,";margin-right:5px;")),
                tags$strong(chain), paste0(" — ",top_chains$n[i]," locations"),
                style="font-size:.78rem;margin-bottom:.2rem;")
            })),
            p(paste0("Total in dataset: ",nrow(state_locs)),
              style="font-size:.7rem;color:#8c7355;font-family:'Space Mono',monospace;")
          )
        )
      )
    })

    output$map_legend <- renderUI({
      div(style="display:flex;flex-wrap:wrap;gap:.6rem;padding:.8rem;
                 background:white;border:1px solid #ede8df;",
        lapply(TOP_CHAINS_LIST, function(chain) {
          col <- get_chain_color(chain)
          tags$span(style="display:flex;align-items:center;gap:4px;font-size:.65rem;
                           font-family:'Space Mono',monospace;",
            tags$span(style=paste0("width:9px;height:9px;border-radius:50%;",
                                   "background:",col,";display:inline-block;")),
            chain)
        })
      )
    })
  })
}
```

---

# Running the App {#run}

```{r run-app}
# From your project root:
shiny::runApp(".")

# Or with options:
shiny::runApp(".", port = 3838, launch.browser = TRUE)
```

## Deploy to shinyapps.io {#deploy}

```{r deploy}
# Install rsconnect once
install.packages("rsconnect")

# Authenticate (get token from shinyapps.io → Account → Tokens)
rsconnect::setAccountInfo(
  name   = "YOUR_ACCOUNT_NAME",
  token  = "YOUR_TOKEN",
  secret = "YOUR_SECRET"
)

# Deploy
rsconnect::deployApp(
  appDir  = ".",
  appName = "fastfood-health-dashboard",
  appFiles = c("app.R", "R/", "data/")
)
```

---

*Generated with R Markdown · Fast Food & Health in America Dashboard*
